---
title: "Data Challenge"
author: "Praveen Burra"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Date : 06/06/2018  

Start Time: 11 PM  

End time : 12:50 AM 


Consider the cleaning of training dataset.
 
```{r}
#Loading the training Data Set
sampledata <- read.csv("C:\\Users\\Satya Praveen\\Desktop\\coding\\data-challenge\\data-challenge\\training_data.csv",sep=',')
head(sampledata) 

# Replacing the missing values with NA
sampledata[sampledata=="?"]<-NA 
head(sampledata)

# dimensions of the training data
dim(sampledata)

#To see how many NA values are there in each column in the given data. 
library(dplyr)
sampledata %>%
select(everything()) %>%  # replace to your needs
summarise_all(funs(sum(is.na(.)))) 

# So we have 7 variables having NA values which are:
# Race, Payer_code,medical_speciality, weight, diag-1,diag-2,diag-3
# Removing the Weight variable which has more than 78844 missing values.
sampledata$weight <- NULL

#Replacing the odd values like V45, E932 with NA in all the three variables(diag-1,diag-2,diag-3)
sampledata$diag_3 <- gsub("V.*",NA, sampledata$diag_3)
sampledata$diag_2 <- gsub("V.*",NA, sampledata$diag_2)
sampledata$diag_1 <- gsub("V.*",NA, sampledata$diag_1)

sampledata$diag_3 <- gsub("E.*",NA, sampledata$diag_3)
sampledata$diag_2 <- gsub("E.*",NA, sampledata$diag_2)
sampledata$diag_1 <- gsub("E.*",NA, sampledata$diag_1)

#FOr diag-3 variable 
#checking whether the variable is numeric or not
is.numeric(sampledata$diag_3) 
#converting to numeric
sampledata$diag_3 <- as.numeric(sampledata$diag_3) 
is.numeric(sampledata$diag_3)
# impute (replace NA values with mean)
sampledata$diag_3[is.na(sampledata$diag_3)] <- mean(sampledata$diag_3, na.rm = T)

#For diag-2
sampledata$diag_2 <- as.numeric(sampledata$diag_2) #converting to numeric
is.numeric(sampledata$diag_2)
#impute 
sampledata$diag_2[is.na(sampledata$diag_2)] <- mean(sampledata$diag_2, na.rm = T)

#For diag-1
sampledata$diag_1 <- as.numeric(sampledata$diag_1) #converting to numeric
is.numeric(sampledata$diag_1)
#impute
sampledata$diag_1[is.na(sampledata$diag_1)] <- mean(sampledata$diag_1, na.rm = T)

#replacing all NA values in the categorical data with None

#### For variable - Race
set.seed(1234)
# Get levels and add "None" level
levels <- levels(sampledata$race)
levels[length(levels) + 1] <- "None"

# refactor Race to include "None" as a factor level and replace NA with "None"
sampledata$race <- factor(sampledata$race, levels = levels)
sampledata$race[is.na(sampledata$race)] <- "None"


#### For variable - payer_code
set.seed(1235)

# Get levels and add "None" level
levels <- levels(sampledata$payer_code)
levels[length(levels) + 1] <- "None"

# refactor payer_code to include "None" as a factor level and replace NA with "None"
sampledata$payer_code <- factor(sampledata$payer_code, levels = levels)
sampledata$payer_code[is.na(sampledata$payer_code)] <- "None"


#### For variable - medical_speciality
set.seed(1236)

# Get levels and add "None"
levels <- levels(sampledata$medical_specialty)
levels[length(levels) + 1] <- "None"

# refactor medical_speciality to include "None" as a factor level and replace NA with "None"
sampledata$medical_specialty <- factor(sampledata$medical_specialty, levels = levels)
sampledata$medical_specialty[is.na(sampledata$medical_specialty)] <- "None"


######## dealing with categorical variables
convert <- c(3,4,5,10,11,22:48)
sampledata[,convert] <- data.frame(apply(sampledata[convert], 2, as.factor))

# checking whether the columns are converted to factor type
is.factor(sampledata[,23])
is.factor(sampledata[,20])

```

Now let us consider the clening of test data.
```{r}
############################### TESTING DATA #################################
testdata = read.csv("C:\\Users\\Satya Praveen\\Desktop\\coding\\data-challenge\\data-challenge\\test_data.csv",sep = ',')
testdata[testdata=="?"]<-NA
head(testdata)

# dimensions of the test data
dim(testdata)

#To see how many NA values are there in each column in the given data. 
library(dplyr)
testdata %>%
  select(everything()) %>%  # replace to your needs
  summarise_all(funs(sum(is.na(.)))) 

# Removing the Weight variable which has more than 19725 missing values.
testdata$weight <- NULL

#Replacing the odd values like V45, E932 and manymore irrelevant numbers with NA in all the three variables
testdata$diag_3 <- gsub("V.*",NA, testdata$diag_3)
testdata$diag_2 <- gsub("V.*",NA, testdata$diag_2)
testdata$diag_1 <- gsub("V.*",NA, testdata$diag_1)

testdata$diag_3 <- gsub("E.*",NA, testdata$diag_3)
testdata$diag_2 <- gsub("E.*",NA, testdata$diag_2)
testdata$diag_1 <- gsub("E.*",NA, testdata$diag_1)

# Imputing the diag-1,2,3 columns missing values with their mean values
#FOr diag-3 variable
#checking whether the variable is numeric or not
is.numeric(testdata$diag_3)
testdata$diag_3 <- as.numeric(testdata$diag_3) 
#converting to numeric
is.numeric(testdata$diag_3)

#impute (replace with mean)
testdata$diag_3[is.na(testdata$diag_3)] <- mean(testdata$diag_3, na.rm = T)

#For diag-2
testdata$diag_2 <- as.numeric(testdata$diag_2) #converting to numeric
is.numeric(testdata$diag_2)
#impute 
testdata$diag_2[is.na(testdata$diag_2)] <- mean(testdata$diag_2, na.rm = T)

#For diag-1
testdata$diag_1 <- as.numeric(testdata$diag_1) #converting to numeric
is.numeric(testdata$diag_1)
#impute
testdata$diag_1[is.na(testdata$diag_1)] <- mean(testdata$diag_1, na.rm = T)

#replacing all NA values in the categorical data with None

#### For variable - Race
set.seed(45)

# Get levels and add "None"
levels <- levels(testdata$race)
levels[length(levels) + 1] <- "None"

# Refactor Race to include "None" as a factor level and replace NA with "None"
testdata$race <- factor(testdata$race, levels = levels)
testdata$race[is.na(testdata$race)] <- "None"

#### For variable - payer_code
set.seed(451)

# Get levels and add "None"
levels <- levels(testdata$payer_code)
levels[length(levels) + 1] <- "None"

# refactor Payer_code to include "None" as a factor level and replace NA with "None"
testdata$payer_code <- factor(testdata$payer_code, levels = levels)
testdata$payer_code[is.na(testdata$payer_code)] <- "None"


#### For variable - medical_speciality
set.seed(452)

# Get levels and add "None"
levels <- levels(testdata$medical_specialty)
levels[length(levels) + 1] <- "None"

# refactor medical_speicality to include "None" as a factor level and replace NA with "None"
testdata$medical_specialty <- factor(testdata$medical_specialty, levels = levels)
testdata$medical_specialty[is.na(testdata$medical_specialty)] <- "None"


######## dealing with categorical variables
convert <- c(3,4,5,10,11,22:48)
testdata[,convert] <- data.frame(apply(testdata[convert], 2, as.factor))

is.factor(testdata[,23]) # checking whether the columns are converted to factor type
is.factor(testdata[,20])

```

We need to handle missing values and categorical features before feeding the data into a machine learning algorithm, because the mathematics underlying most machine learning models assumes that the data is numerical and contains no missing values. 

```{r}
########################### MODELLING #######################
#install.packages("h2o")
library(h2o)

#To launch the H2O cluster
localH2O <- h2o.init(nthreads = -1)

# loading the training data into H2o environment from R
train.h2o <- as.h2o(sampledata)

test.h2o <- as.h2o(testdata)

colnames(train.h2o)


### GBM(Gradient Boosting) Model
y.dep <- 49 #deppendent variable
# Removing the encounter id and patient number from the dependent variables as they don't add any extra information for training the model.
x.indep <- c(3:48) #independent variables 

#
#GBM model on the training data set in H2o environment
gbm.model <- h2o.gbm(y=y.dep, x=x.indep, training_frame = train.h2o, ntrees = 1000, max_depth = 4, learn_rate = 0.01, seed = 1122)

# predicting the readmission results for the test data set 
predict.gbm <- as.data.frame(h2o.predict(gbm.model, test.h2o))

summary(predict.gbm)
# Shutting down the h2o environment
h2o.shutdown(prompt = FALSE)
```

Random Forests and othe models are also used but they are taking longer execution times because of large dataset. For reference we can see the code below.

```{r}
#### test for selecting the important variables
#install.packages("Boruta")
#library(Boruta)
#set.seed(123)
#boruta_output <- Boruta(readmitted ~ ., data=sampledata, doTrace=2)  # perform Boruta search
#print(boruta.train)


#install.packages("randomForest")
#library(randomForest)
#model_rf<-randomForest(readmitted ~ ., data = sampledata[,48:49])
#These models and methods are taking too much time to process. Hence I have used H2o Environment linked with R for dealing with large datasets.

```


Final Predicted values of readmission of diabetes patients for the test dataset. Since the final outcomes of the testdata file are not given here, the confusion matrix and the accuracy scores of the model on the testdata couldn't be calculated.
```{r}
# Final Output predictions dataframe containing the row identifiers(encounter_id and patient number)
Final_Output <- data.frame(encounter_id = testdata$encounter_id, readmitted = predict.gbm$predict)

#  Exporting the csv file and saving it.
write.csv(Final_Output, file = "Burra_Praveen.csv", row.names = F)

```

